---
title: "sesameData User Guide"
shorttitle: "sesameData guide"
package: sesameData
output: rmarkdown::html_vignette
fig_width: 8
fig_height: 6
vignette: >
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteIndexEntry{sesameData User Guide}
    %\VignetteEncoding{UTF-8}
---

# Installation

From Bioconductor
```{r, eval=FALSE}
if (!requireNamespace("BiocManager", quietly=TRUE))
    install.packages("BiocManager")
BiocManager::install("sesameData")
```

Development version can be installed from github
```{r, eval=FALSE}
library(devtools)
install_github('zwdzwd/sesameData')
```

# Usage

`sesameData` package provides associated data for sesame package. This includes
example data for testing and instructional purpose, as we ll as probe annotation
for different Infinium platforms.

## Load sesameData
```{r, message=FALSE, warning=FALSE}
library(sesameData)
```

## Caching data on local disk

Each datum is accessible through the `sesameDataGet` interface. It should be
noted that all data must be pre-cached to local disk before they can be used.
This design is to prevent conflict in annotation data caching and remove
internet dependency. Caching needs only be done once per sesame/sesameData
installation. One can cache data for a specific assay platform, say HM27 by
```{r}
sesameDataCache("HM27")
```
or choose to cache all sesame data, by
```{r, eval=FALSE}
sesameDataCacheAll()
```

Once a data object is loaded, it is stored to a tempoary cache, so that the data
doesn't need to be retrieved again next time we call `sesameDataGet`. This
design is meant to speeed up the run time.

## Retrieval data

For example, the annotation for HM27 can be retrieved with the title: 
```{r}
HM27.hg38.manifest = sesameDataGet('HM27.hg38.manifest')
```

## List available data

Titles of all the available data can be shown with:
```{r}
sesameDataList()
```

## Download .tsv.gz from the annotation website
```{r}
mft <- sesameDataGetAnno("HM27/HM27.hg19.manifest.tsv.gz")
```

## Download RDS files from the annotation website
```{r}
annoS <- sesameDataGetAnno("EPIC/EPIC.hg19.typeI_overlap_b151.rds")
```

## Download annotation file to disk
```{r}
sesameDataDownload("3999492009_R01C01_Grn.idat", dest_dir = tempdir())
```

```{r}
sessionInfo()
```