---
title: "sesameData User Guide"
shorttitle: "sesameData guide"
package: sesameData
output: rmarkdown::html_vignette
fig_width: 8
fig_height: 6
vignette: >
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteIndexEntry{sesameData User Guide}
    %\VignetteEncoding{UTF-8}
---

# Installation

From Bioconductor
```{r, eval=FALSE}
if (!requireNamespace("BiocManager", quietly=TRUE))
    install.packages("BiocManager")
BiocManager::install("sesameData")
```

Development version can be installed from github
```{r, eval=FALSE}
library(devtools)
install_github('zwdzwd/sesameData')
```

# Usage

`sesameData` package provides associated data for sesame package. It also 
maintains probe annotation for different Infinium platforms.

## Load sesameData
```{r, message=FALSE, warning=FALSE}
library(sesameData)
```

## Caching data on local disk

Each datum is accessible through the `sesameDataGet` interface which allows
package level in-memory caching. It should be noted that all data must be
pre-cached to local disk before they can be used. This design is to prevent
conflict in annotation data caching and remove internet dependency. Caching
needs only be done once per sesame/sesameData installation. One can cache
data for a specific assay platform, say HM27 by
```{r}
sesameDataCache("HM27")
```
or choose to cache all sesame data, by
```{r, eval=FALSE}
sesameDataCacheAll()
```

## Retrieval data

Then data can be retrieved with the title: 
```{r}
HM27.hg38.manifest = sesameDataGet('HM27.hg38.manifest')
```

## List available data


Titles of all the available data can be shown with:
```{r}
sesameDataList()
```


